---
title: "Santa Barbara Urchins"
author: "Eli Schulman"
date: "10/23/2021"
output: html_document
---

Overview:

***Data Citation: *** Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b. 

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# attach packages
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(lubridate)
```

```{r}
urchins <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()
```

### Purple vs Red urchins (distributions of sizes)

```{r}
ggplot(urchins, aes(x = size)) +
  geom_histogram(aes(fill = common_name), bins = 10) +
  labs(title = "Purple and Red Urchin Size Distribution",
       x = "Diameter (cm)",
       y = "# of Urchins",
       fill = "Urchin Color") +
# change legends color
  scale_fill_manual(values = c("blueviolet","red3")) +
# center the plot title
  theme(plot.title = element_text(hjust = .5)) +
  facet_wrap(~ common_name)
```

### Descriptive statistics for Urchins

```{r}
urchins_stats <- urchins %>% 
  group_by(common_name) %>% 
  summarize(mean = signif(mean(size), 2),
            median = signif(median(size), 2),
            sd = signif(sd(size), 2),
            sample_size = n())

kable(urchins_stats, 
      col.names = c('Species', 'Mean (cm)', 'Median (cm)', 'Standard Deviation (cm)', 'Sample Size'),
      caption = "Summary Statistics of Urchin Species Body Diameters") %>% 
  kable_styling(full_width = FALSE)
```

### Takeaway from urchin species distribution and stats summary: 
#### The red urchins are more normaly distributed, and have a larger mean diameter relative to the purple urchins (6.78 cm vs 4.47 cm). Additionally, the red urchins have a larger spread in observed diameters compared to their mean.

### Purple urchins population sampling by site by year.

```{r}
purple_urchins <- urchins %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date)) %>% 
  mutate(site_full = case_when(
    site == "NAPL" ~ "Naples Reef",
    site == "AQUE" ~ "Arroyo Quenado Reef",
    site == 'CARP' ~ 'Carpinteria Reef',
    site == 'MOHK' ~ 'Mohawk Reef',
    site == 'IVEE' ~ 'Isla Vista Reef',
    TRUE ~ "NA"
  ))

# Create summary table
purple_urchins_trend <- purple_urchins %>% 
  group_by(year, site_full) %>% 
  summarize(pop = n())
```

### Graph Purple Urchin trends by site by year

```{r}
ggplot(purple_urchins_trend, aes(x = year, y = pop)) +
  geom_line(aes(color = site_full)) +
# apply trend line
  geom_smooth(method = lm, se = FALSE, color = 'Black') +
  labs(title = "Purple Urchin Decline 2008-2020",
       x = "Year",
       y = "Urchin Observations",
       color = "Site") +
  theme_minimal() +
# center plot title
  theme(plot.title = element_text(hjust = .5)) 
```


Data Interpretation: Population sampling of Purple Urchins indicate that, at every site in the study, populations have declined between 2008 and 2020. The sampling data could include errors resulting from differences in sampling methodology. Additionally, kelp treatment at the sites may have an impact on observed population sizes.

QUESTIONS:
1. half years for purple urchin trends

