---
title: "Naples Reef (Jan 2010) Purple Urchins"
author: "Eli Schulman"
date: "10/24/2021"
output: html_document
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# attach packages
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(lubridate)
```

```{r}
urchins_purp <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names() %>%  
# convert date column using lubridate, then create a year variable
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date)) %>% 
# filter for Naples, purple urchins, and 2010 data
  filter(common_name == "Purple Urchin", 
         site == 'NAPL',
         year == 2010)
```

### Exploratory graphics (distribution histogram and Q-Q plot)

```{r}
# histogram of purple urchins to show distribution of sizes at naples between control and annual groups
ggplot(urchins_purp, aes(x = size)) +
  geom_histogram(aes(fill = treatment), bins = 10)

# Q-Q plot to show "correlation"?????
ggplot(urchins_purp, aes(sample = size)) +
  geom_qq(aes(color = treatment), shape = 5, size = 4)

```

#### Distribution Interpretation: The samples are normally distributed, and exhibit a slight positive skew. Additionaly, removing the giant kelp each winter (ANNUAL treatment) results in smaller sample sizes of purple urchins at Naples Reef.

### Summary Statistics for purple urchins at Naples Reef (Control vs. Annual)

```{r}
urchins_purp

```





